A GPU Implementation of Capon for an Active Sonar System

### 2012-05-03 ECUA2012/gpu_capon_ecua2012.tex
la2pdf commit - 2012:05:03	21:58:21	(Thu 3.May)

@@

Consider an $M$ element phased array. Let the sample recorded by the $m$'th channel at time instant $n$ be represented as $x_m[n]$, and the accompanying weights and delays be  $w_m[n]$ and $\Delta_m[n]$. The beamformer output $z[n]$ is then given as
\begin{align}
z[n] = \w\H[n]\x[n] = \bmat{w_0[n]\\w_1[n]\\\vdots\\w_{M-1}[n]}^H \bmat{x_0[n-\Delta_0]\\x_1[n-\Delta_1]\\\vdots\\x_{M-1}[n-\Delta_{M-1}]}.\label{z}
\end{align}
The \gls{MVDR} beamformer \cite{Capon1969} finds the set of weights that minimizes the output power, $\min E\{|z[n]|^2\}$, while ensuring unity gain in the look direction. The solution to this optimization problem is given as
\begin{gather}
\vec w[n] = \frac{\Ri[n]\1}{\1\T\Ri[n]\1},\label{weights}
\end{gather}
where $\1$ is a row vector of ones, and $\R$ is the spatial covariance matrix. Since $\R$ is unknown, it must be estimated. We will do this by performing some degree of spatial averaging to avoid signal cancellation, and temporal averaging to achieve \gls{DAS}-like speckle statistics. We start with the former, by segmenting the array into $M-L-1$ subarrays of length $L$, where all but one element overlap. The data vector $x_l[n]$ from subarray $l$ is given as
\begin{gather}
\x_l[n] = \bmat{x_l[n-\Delta_l] & x_{l+1}[n-\Delta_l] & \dots & x_{l+L-1}[n-\Delta_{l+L-1}]}\T,
\end{gather}
and the spatially averaged sample covariance matrix, $\tilde R$, is given as
\begin{gather}
\tilde\R[n] = \frac{1}{M-L+1} \sumb{l=0}{M-L+1} \x_l[n]\x_l\H[n].\label{spatialR}
\end{gather}
Finally, we take the average of $2N_\text{avg}+1$ temporal samples to form the final sample covariance matrix $\eR$,
\begin{gather}
\eR[n] = \frac{1}{2N_\text{avg}+1} \sumb{n'=n-N_\text{avg}}{n+N_\text{avg}} \tilde\R[n'].
\end{gather}
Finally, to make sure that $\eR$ is well conditioned, we add a fraction $d$ of the total power of $\eR$ to its diagonal\cite{Synnevag2007},
\begin{align}
\eR[n] = \eR[n] + \I \cdot \frac{d}{L} \cdot \tr\{\eR[n]\}.\label{finalR}
\end{align}
To sum it up, for each range line, and for each sample $n$ in this line we must
\begin{enumerate}
\item Compute the spatial covariance matrix $\eR[n]$ (equation \ref{spatialR}-\ref{finalR}).
\item Compute $\eRi[n]\1$ (from equation \ref{weights}).
\item Insert the result from step (2) into equation \ref{weights} to find the weights.
\item Insert the result from step (3) into equation \ref{z} to acquire the pixel value.
\end{enumerate}

@@

Massively parallel hardware such as \glspl{GPU} have a potential for accelerating many signal processing tasks, but fully exploiting their potential is not always trivial. Modern \gls{GPU} chips contain several hundred processing cores, but the amount of memory available locally for each core is very limited. There is also a considerable latency involved when accessing on-board memory, so assessing various memory access schemes is crucial to achieve optimal performance.

Hence, the programming strategy should be to decompose the algorithm into as many threads as possible, each being computationally intense and be light on memory access and consumption. Consecutive threads should also access consecutive memory regions, to enable coalesced data transfers, and have the same line of execution.

>>>>

Massively parallel hardware such as \glspl{GPU} have a potential for accelerating many signal processing tasks, and the \gls{MVDR} beamformer is no exception. Each pixel in the image can be processed independently, and

Modern \gls{GPU} chips contain several hundred processing cores, but the amount of memory available locally for each core is very limited.\todo{repetition} There is also a considerable latency involved when accessing on-board memory, but these memory transfers are asynchronous and may be hidden.

@@

Consider an $M$ element phased array. Let the sample recorded by the $m$'th channel at time instant $n$ be represented as $x_m[n]$, and the accompanying weights and delays be  $w_m[n]$ and $\Delta_m[n]$. The beamformer output $z[n]$ is then given as
\begin{align}
z[n] = \w\H[n]\x[n] = \bmat{w_0[n]\\w_1[n]\\\vdots\\w_{M-1}[n]}^H \bmat{x_0[n-\Delta_0]\\x_1[n-\Delta_1]\\\vdots\\x_{M-1}[n-\Delta_{M-1}]}.\label{z}
\end{align}
The \gls{MVDR} beamformer \cite{Capon1969} finds the set of weights that minimizes the output power, $\min E\{|z[n]|^2\}$, while ensuring unity gain in the look direction. The solution to this optimization problem is given as
\begin{gather}
\vec w[n] = \frac{\Ri[n]\1}{\1\T\Ri[n]\1},\label{weights}
\end{gather}
where $\1$ is a row vector of ones, and $\R$ is the spatial covariance matrix. Since $\R$ is unknown, it must be estimated. We will do this by performing some degree of spatial averaging to avoid signal cancellation, and temporal averaging to achieve \gls{DAS}-like speckle statistics. We start with the former, by segmenting the array into $M-L-1$ subarrays of length $L$, where all but one element overlap. The data vector $x_l[n]$ from subarray $l$ is given as
\begin{gather}
\x_l[n] = \bmat{x_l[n-\Delta_l] & x_{l+1}[n-\Delta_l] & \dots & x_{l+L-1}[n-\Delta_{l+L-1}]}\T,
\end{gather}
and the spatially averaged sample covariance matrix, $\tilde R$, is given as
\begin{gather}
\tilde\R[n] = \frac{1}{M-L+1} \sumb{l=0}{M-L+1} \x_l[n]\x_l\H[n].\label{spatialR}
\end{gather}
Finally, we take the average of $2N_\text{avg}+1$ temporal samples to form the final sample covariance matrix $\eR$,
\begin{gather}
\eR[n] = \frac{1}{2N_\text{avg}+1} \sumb{n'=n-N_\text{avg}}{n+N_\text{avg}} \tilde\R[n'].
\end{gather}
Finally, to make sure that $\eR$ is well conditioned, we add a fraction $d$ of the total power of $\eR$ to its diagonal\cite{Synnevag2007},
\begin{align}
\eR[n] = \eR[n] + \I \cdot \frac{d}{L} \cdot \tr\{\eR[n]\}.\label{finalR}
\end{align}

@@

By the methods just presented, we observe that for each range line, and for each sample $n$, the following steps must be carried out:
\begin{enumerate}
\item \emph{Compute the spatial covariance matrix} $\eR[n]$ (equation \ref{spatialR}-\ref{finalR}).
\item Compute $\eRi[n]\1$ (from equation \ref{weights}).
\item Substitute the solution for $\eRi[n]\1$ into equation \ref{weights} to find the weights.
\item Apply the weights to each data vector $\x_l[n]$, and sum them all up (as in equation \ref{z}) to acquire the pixel value $z[n]$.
\end{enumerate}

Massively parallel hardware such as \glspl{GPU} have a potential for accelerating many signal processing tasks, but fully exploiting their potential is not always trivial. Modern \gls{GPU} chips contain several hundred processing cores, but the amount of memory available locally for each core is very limited.\todo{repetition} There is also a considerable latency involved when accessing on-board memory, but these memory transfers are asynchronous and may be hidden.

decompose the algorithm into as many threads as possible, each being computationally intense and be light on memory access and consumption. Consecutive threads should also access consecutive memory regions, to enable coalesced data transfers, and have the same line of execution.

How may this be decomposed? Already decomposed at pixel-level.

We implement the 
Our implementation of th

When
